As documented in this chapter, several idioms that we we can count on when using containers are broken by views. Here is a quick summary you should always take into account when using generic code for both containers and views:

\begin{itemize}
\item
You might not be able to iterate over the elements of a standard view when it is const.

As a consequence, generic code for all kinds of ranges (containers and views) has to declare the parameters as universal/forwarding references.

However, do not use universal/forwarding references when you have concurrent iterations. In that case, const is your friend.

\item
Standard views on lvalue ranges do not propagate constness.

This means that declaring such a view as const does not declare the elements to be const.

\item
Concurrent iterations over standard views might cause data races (a runtime error due to undefined behavior) even if they only read.

\item
Reading iterations might affect later functional behavior or even invalidate later iterations.

Use standard views ad hoc.

\item
Copying a view might create a view with different states and behavior than the source view.

Avoid copying standard views.

\item
cbegin() and cend() might not make elements const.

This will partially be fixed with C++23 by providing cbegin() and cend() member functions and fixing std::ranges::cbegin() and std::ranges::cend(). Unfortunately, std::cbegin() and std::cend() will still be broken.

\item
Type const\_iterator is usually not available.

\item
For C++23, the following breakage is planned: for some standard views, declaring the elements as const might have no effect. You might be able to modify the members of a const element of a view.
\end{itemize}

This means that a standard view is not always a pure subset that restricts or deals with the elements of a range; it might provide options and operations that are not allowed for when using the range as a whole instead.

As a consequence, use views ad hoc and with care and forget about any temporary constness. Honestly, you might consider avoiding standard views and using something with a safer design instead.